---
title: "DAIS Health"
output: html_document
date: "2023-04-21"
---




```{r}
demo <- read_csv("~/Desktop/DAIS_PROJ/demo_DEMO.csv")%>%
  select(2, 5, 6, 15)
```

The partial dataframe from the original demography csv file: Left-join beginner.



### Below chunks are sample codes I used in editing alcohol csv data.

```{r}
alc <- read_csv("~/Downloads/alc.csv")%>%
  mutate_at(vars(`DR2IALCO`), funs(round(., 2)))%>%
  drop_na()
```

Reads csv file "alc.csv" then rounds all the value under the column "DR2IALCO." Please do make sure you round the integer minimum values to make them 0. I think it was like 5.40E-79. 

```{r}
alc<-alc[alc$DR1IALCO != 0, ]%>%
  drop_na()
```

Drops rows with null value under variable (column) DR1IALCO. 

```{r}
df1[is.na(df1)] <- 0
```

Replaces every null value in the dataframe df1 with 0. 

```{r}
liver <- left_join(demo, bmx, by = 'SEQN')%>%
  drop_na(`BMXWT`, `BMXBMI`, `LBXTR`)
```

"demo" will be your base dataframe and "bmx" will join the base by the axis column "SEQN." The rows will be preserved despite mismatch. Previous null drop operation via dplyr pipeline (%>%) has been applied on the specified variables (columns "`BMXWT`, `BMXBMI`, `LBXTR`".)

  
```{r}
%>%
  mutate(`...1` = NULL)
```

Use code above if you want to drop the column `...1`

```{r}
write.csv(liver3, "~/Desktop/DAIS_PROJ/liveranal.csv")
```

Writes csv file in the designated destination.

